---
title: "Physical Excercise Form Classification Model"
author: "Forest"
date: "August 16, 2014"
output: html_document
---

##Abstract
In this R Markdown report, we examine the data collected from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. Our goal is to classify whether they perform barbell lifts correctly and incorrectly in 5 different ways.


Here are the R packages we will need for this analysis:
```{r library, message=FALSE, warning=FALSE}
library(caret)
library(dplyr)
library(corrplot)
```

##Data Import and Cleaning
```{r load}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

There are a lot of variables and some of them have large amounts of missing values. We want to first filter out those unhelpful one and move to feature selection from there.

```{r clean, warning=FALSE}
#save classe variable
trainingClasse <- training$classe
#find all factor variables
factorCols <- sapply(training, is.factor)
#force them to numeric
toNumeric <- function(x) { as.numeric(as.character(x)) }
training <- mutate_each_q(training, funs(toNumeric), colnames(training)[factorCols])
#filter all columns with NAs
badColumns <- sapply(training, function(x){ sum(is.na(x)) > 0 }) 
goodTraining <- training[, !badColumns] %>% mutate(classe = trainingClasse)
#filter out index and time columns
goodTraining <- select(goodTraining, -X, -contains("time"), -num_window)
```

### Model Fitting
We will now configure 
```{r train}
set.seed(1431)
#set up cross validation
cvControl <- trainControl(method="cv")
#set up number of variables used in growing trees
rfGrid <- data.frame(mtry=10:10)
rfFit <- train(classe ~ ., goodTraining, tuneGrid = rfGrid, trControl = cvControl)
```

### Model testing